---
title: "Final Proj Code"
output: html_notebook
---

```{r}
# Importing Libraries 
library(datarium)
```

```{r}
#Importing Dataset 
data("headache")
library(readxl)
headachelong <- read_excel("~/Desktop/ANOVA/headachelong.xlsx")
View(headachelong)
```

# Descriptive Statistics 
```{r}

```

## For quantitative variables:
  Numerical Measures
  Mean
  Std Dev
  Skew
  Kurtosis
  Boxplots & Outliers
  Equal Variance tests (& their limitations)
  Normality Tests: qq plots

## For Categorical variables:
  Frequency charts or tally charts
  Means plots & Interaction plots
```{r}
ggline(headache, x = "id", y = "pain_score", col="risk", add=c("mean_se"))
```
  
  Discussion of what all of this means

# Overview of statistical analyses performed
  What tests: Overall, post hocs, contrasts
  Null for each in terms of the question under study and variables
  Assumptions needed for each test (e.g., normality, equal variances, sphericity, non-interaction of     any blocks (or subjects) with other factors, etc.
  
# Empirical power of your tests. 
  What assumptions did you make in calculating this power?
  If power is less than 90%, what sample size is needed?

