---
title: "Final Proj Code"
output: html_notebook
---

```{r}
# Importing Libraries 
library(readxl)
library(ez)
library(ggplot2)
library(multcomp)
library(nlme)
library(pastecs)
library(reshape2)
library(haven)
library(dplyr)
library(ggpubr)
library(rstatix)
library(mosaic)
library(car)
library(moments)
```

```{r}
#Importing Dataset 
headachelong<-read_excel("headachelong.xlsx")
```

## For quantitative variables:
  Numerical Measures
  Mean
  Std Dev
  Skew
  Kurtosis
  Boxplots & Outliers
  Equal Variance tests (& their limitations)
  Normality Tests: qq plots
  
# Descriptive Statistics 
```{r}
# Quantitative Variable - Pain Score Descriptives 
painscore<-headachelong$painscore #saving painscore to a variable 
descriptives<-favstats(painscore)
descriptives$mean # shows mean 
descriptives$sd # shows sd
```

```{r}
# Quantitative Variable -  Pain Score Descriptives 
skewness(painscore)
kurtosis(painscore)
ks.test(painscore,"pnorm")
hist(painscore)
jarque.test(painscore)
```
Skewness = 0.58 > 0, deviates from normality - right skewed, skew is not that bad because it is less than 1 
Kurtosis = 3.52 > 3 , thinner tailed than normal, leptokurtic, not too bad. 
Ks.test = pval < 0.05, reject null of normality 
Jarque Test = pval > 0.05 fail to reject that data has same skew and kurtosis that a normal with the same mean and sd has. 

```{r}
#Quantitative Variable - Pain Score Descriptives 

#Boxplots 
gf_boxplot(painscore~gender,data = headachelong) 
gf_boxplot(painscore~trmt,data=headachelong)
gf_boxplot(painscore~risk,data=headachelong)

#Equal Variance Test 
leveneTest(painscore~gender,data=headachelong)
leveneTest(painscore~trmt,data=headachelong)
leveneTest(painscore~risk,data=headachelong)

#Normality Test: qqplot 
gf_qq(~painscore,data=headachelong) %>% gf_qqline()
```
In the boxplots for each, there seems to be outliers. 


## For Categorical variables:
  Frequency charts or tally charts
  Means plots & Interaction plots
```{r}
#Tally 
tally(~gender,data = headachelong) #gender
tally(~trmt,data = headachelong) #treatment
tally(~risk,data = headachelong) #risk 

#Interaction plots 
ggline(data = headachelong, x = "gender", y = "painscore", col = "risk", add = c("mean_se"))
ggline(data = headachelong, x = "trmt", y = "painscore", col = "risk", add = c("mean_se"))
ggline(data = headachelong, x = "trmt", y = "painscore", col = "gender", add = c("mean_se"))

#Creating subset for threeway interaction
painscore.m<-subset(headachelong,gender == "male")
ggline(data = painscore.m, x = "trmt", y = "painscore", col = "risk", add = c("mean_se"),main = "Male")
painscore.f<-subset(headachelong,gender == "female")
ggline(data = painscore.f, x = "trmt", y = "painscore", col = "risk", add = c("mean_se"),main = "Female")
painscore.x<-subset(headachelong,trmt == "X")
ggline(data = painscore.x, x = "gender", y = "painscore", col = "risk", add = c("mean_se"), main = "Treatment X")
painscore.y<-subset(headachelong,trmt == "Y")
ggline(data = painscore.y, x = "gender", y = "painscore", col = "risk", add = c("mean_se"), main = "Treatment Y")
painscore.z<-subset(headachelong,trmt == "Z")
ggline(data = painscore.z, x = "gender", y = "painscore", col = "risk", add = c("mean_se"), main = "Treatment Z")
painscore.low<-subset(headachelong,risk  == "low")
ggline(data = painscore.low, x = "gender", y = "painscore", col = "trmt", add = c("mean_se"), main = "Low Risk")
painscore.high<-subset(headachelong,risk  == "high")
ggline(data = painscore.high, x = "gender", y = "painscore", col = "trmt", add = c("mean_se"), main = "High Risk")
```
  
# Overview of statistical analyses performed
  What tests: Overall, post hocs, contrasts
  Null for each in terms of the question under study and variables
  Assumptions needed for each test (e.g., normality, equal variances, sphericity, non-interaction of     any blocks (or subjects) with other factors, etc.


Testing a null of no difference in mean painscores
**ORTHOGONAL**
Hypothesis 1: Treatment Y is better than X. 
Hypothesis 2: Treatment Z is better than X and Y 
```{r}
# Creating Contrasts - Treatments *ORTHOGONAL*
headachelong$trmt<-as.factor(headachelong$trmt)
YvsX<-c(0,-1,1)
ZvsOthers<-c(-2,1,1)

#headachelong$risk<-as.factor(headachelong$risk)
#lowXvshighX<-c()
#lowYvshighY<-c()
#lowZvshighZ<-c()

contrasts(headachelong$trmt)<-cbind(YvsX,ZvsOthers)
headachelong$trmt

#contrasts(headachelong$risk)<-cbind(lowXvshighX,lowYvshighY,lowZvshighZ)
#headachelong$risk
```

```{r}
# ANOVA Model
headache.mod<-lme(painscore~trmt+gender+risk+trmt:gender+trmt:risk+gender:risk+trmt:gender:risk,random= ~1|id/gender/risk,data=headachelong,method="ML")

#headache.mod<-ezANOVA(headachelong,dv = .(painscore),wid=.(id),within=.(risk,trmt),type = 3, detailed=TRUE)
#headache.mod

```

```{r}
# Post Hocs
pairwise.t.test(headachelong$painscore,headachelong$trmt,paired = TRUE, p.adjust.method = "bonferroni")
```

Testing a null of no difference in mean painscores
**NON-ORTHOGONAL**
Hypothesis 1: 
Hypothesis 2: 
```{r}
# Creating Contrasts - Treatments *NON-ORTHOGONAL*
headachelong$trmt<-as.factor(headachelong$trmt)
XvsY<-c(1,0,0)
YvsZ<-c(0,0,1)

contrasts(headachelong$trmt)<-cbind(XvsY,YvsZ)
headachelong$trmt
```

```{r}
# ANOVA Model
headache.mod.nonorthogonal<-lme(painscore~trmt+gender+risk+trmt:gender+trmt:risk+gender:risk+trmt:gender:risk,random= ~1|id/gender/risk,data=headachelong,method="ML")
summary(headache.mod)
```

```{r}
# Post Hocs
pairwise.t.test(headachelong$painscore,headachelong$trmt,paired = TRUE, p.adjust.method = "bonferroni")
```

# Analysis of residuals
What assumptions about your data are you looking to verify?
Which plots to use for which assumption. Why?
Your results.
How valid is your analysis?
```{r}
#Analysis of Residuals - Orthogonal 
par(mfrow = c(2,2))
plot(headache.mod.nonorthogonal,which=5)
plot(headache.mod.nonorthogonal,which = 1)
plot(headache.mod.nonorthogonal,which = 2)
headachelong$id<-c(1:72)
plot(residuals(headache.mod.nonorthogonal)~id, main = "Residuals vs Exp.Unit",font.main = 1, data = headachelong)
```


# Empirical power of your tests. 
  What assumptions did you make in calculating this power?
  If power is less than 90%, what sample size is needed?
```{r}
#Empirical Power of Tests 

set.seed(100)

nreps=1000

pval<-numeric(nreps)

result<-numeric(nreps)

donotreject=0

for (i in 1:nreps){
y1<-rnorm(10,60,2)
y2<-rnorm(10,65,15) #means are NOT same, so NULL is false
t<-t.test(y1,y2,var.equal=TRUE)
pval[i]<-t$p.value
if (pval[i]>=0.05){
  
 ##result[i]= pval[i]
 donotreject=donotreject+1
} else donotreject=donotreject # result[i]=1


}
donotreject
type2=donotreject/nreps
type2
power=1-type2
power
```


